---
title: "Break-Study-Project"
format: html
execute: 
  echo: false
  warning: false
  message: false
embed-resources: true
  
---

```{r}
#| label: setup
#| include: false
library(tidyverse)
library(haven)
library(janitor)
```
## Data Description

The data for this study was collected by PSY 321 Fall 2024 students and utilized a randomized group design.

The present study’s research question was: Does spending class break “as usual,” in nature, or engaged in movement impact affect and stress? 
Method: The present study used a group-randomized design to assign twenty-one university classes into three conditions: (1) break as usual, (2) outdoor break, and (3) indoor movement break. From these classes, participants included 374 students who completed an anonymous pre-break survey, 10-minute break, and then post-break survey measuring stress levels and positive/negative affect (20-item Positive and Negative Affect Schedule). Facilitation of break conditions was standardized using slides and a script. 

Results: We used a mixed within-between subjects ANOVA to assess the impact of the conditions on participants’ affect and stress scores, across time (pre- and post-intervention). There was a statistically significant increase in positive affect, decrease in negative affect, and decrease in stress when students across all break conditions (p < .001); effect sizes moderate to large (partial eta squared).

Conclusions: Results indicated a significant and meaningful difference in stress and affect across conditions. Our hypothesis – that outdoor and movement breaks would be more beneficial than “break as usual” (i.e., often on screens) – was not supported by the data. Our research provides preliminary evidence for the use of unstructured breaks as a strategy for improving student wellness in the classroom.

## Data Cleaning

In order to assess the data, I needed to import my data from SPSS. I achieved that using the haven package and then I had to make sure that the variables were in the correct format. Adjustments needed to be made based on the names of variables and organization of observations in the dataset. I also made sure that that I did not have missing values in my observations. 

```{r}
#| label: data-cleaning
# this should be hidden from view and not seen in the website. It should include your data cleaning and you should export the clean data.
CBI24 <- read_sav("data/PSY321 Class Break Intervention Study FA24_11.8.24_JL_DataClean3.sav")

CBI24_2 <- read.csv("data/PSY321 Class Break Intervention Study FA24_11.8.24_JL_DataClean3_1.csv")

CBI24_2 %>% 
  clean_names()
```

```{r}
#| label: read-data
#read in your clean data here
iris <- read_csv("data/clean-iris.csv")
cares <- read_csv("data/clean-cars.csv")
```

## Visualzation 1

Here is a description of the first visualization. I want to describe what the visualization shows to a general audience. I do not want to see any code or technical terms. I want to be able to read this and understand what the visualization shows. I also want to see a caption and alt text for the visualization. The caption should be a short description of the visualization and the alt text should be a longer description of the visualization. The alt text should be descriptive enough that someone who is blind could understand what the visualization shows.

```{r}
#| label: viz-1
#| fig-cap: "This is a caption for the first visualization"
#| fig-alt: "This is an alternative text for the first visualization"
#| fig-width: 6
#| fig-height: 4
#| fig-align: center

#create plot of iris dataset
ggplot(iris, aes(x = Sepal.Length, y = Sepal.Width)) +
  geom_point() +
  labs(title = "Iris Dataset", 
       x = "Sepal Length", 
       y = "Sepal Width") +
  theme_minimal()
```

## Visualization 2

Here is a description of the second visualization. I want to describe what the visualization shows to a general audience. I do not want to see any code or technical terms. I want to be able to read this and understand what the visualization shows. I also want to see a caption and alt text for the visualization. The caption should be a short description of the visualization and the alt text should be a longer description of the visualization. The alt text should be descriptive enough that someone who is blind could understand what the visualization shows.

```{r}
#| label: viz-2
#| fig-cap: "This is a caption for the second visualization"
#| fig-alt: "This is an alternative text for the second visualization"
#| fig-width: 6
#| fig-height: 4
#| fig-align: center

#create plot of cars data set
ggplot(cars, aes(x = speed, y = dist)) +
  geom_point() +
  labs(title = "Cars Dataset", 
       x = "Speed", 
       y = "Distance") +
  theme_minimal()
```
