---
title: "Break-Study-Project"
format: html
execute: 
  echo: false
  warning: false
  message: false
embed-resources: true
  
---

```{r}
#| label: setup
#| include: false
library(tidyverse)
library(haven)
library(janitor)
library(data.table)
```
## Data Description

The data for this study was collected by PSY 321 Fall 2024 students and utilized a randomized group design.

The present study’s research question was: Does spending class break “as usual,” in nature, or engaged in movement impact affect and stress? 
Method: The present study used a group-randomized design to assign twenty-one university classes into three conditions: (1) break as usual, (2) outdoor break, and (3) indoor movement break. From these classes, participants included 374 students who completed an anonymous pre-break survey, 10-minute break, and then post-break survey measuring stress levels and positive/negative affect (20-item Positive and Negative Affect Schedule). Facilitation of break conditions was standardized using slides and a script. 

Results: We used a mixed within-between subjects ANOVA to assess the impact of the conditions on participants’ affect and stress scores, across time (pre- and post-intervention). There was a statistically significant increase in positive affect, decrease in negative affect, and decrease in stress when students across all break conditions (p < .001); effect sizes moderate to large (partial eta squared).

Conclusions: Results indicated a significant and meaningful difference in stress and affect across conditions. Our hypothesis – that outdoor and movement breaks would be more beneficial than “break as usual” (i.e., often on screens) – was not supported by the data. Our research provides preliminary evidence for the use of unstructured breaks as a strategy for improving student wellness in the classroom.

## Data Cleaning

In order to assess the data, I needed to import my data from SPSS. I achieved that using the haven package and then I had to make sure that the variables were in the correct format. Adjustments needed to be made based on the names of variables and organization of observations in the dataset. I also made sure that that I did not have missing values in my observations. To create a fidelity check on what activity students spent their breaks on by break conditions, I had to group them by condition, and summarize the data to get an outcome of what each person did based on condition.

```{r}
#| label: data-cleaning
# this should be hidden from view and not seen in the website. It should include your data cleaning and you should export the clean data.

# How do you usually spend your break?
#note this code is to demonstrate a fidelity check with original SPSS data
break_study <- read_sav("data/data_raw/PSY321 Class Break Intervention Study FA24_11.8.24_JL_DataClean3.sav")

cbi24 <- read.csv("data/data_raw/PSY321 Class Break Intervention Study FA24_11.8.24_JL_DataClean3_1.csv")

ba_1 <- cbi24 %>% 
  clean_names() %>%
  group_by(break_condition) %>%
  summarize(bathroom = sum(break_condition_3, na.rm = TRUE),
          screenSchool = sum(break_condition_5, na.rm = TRUE),
          screenSocial = sum(break_condition_6, na.rm = TRUE),
          screenRelax = sum(break_condition_7, na.rm = TRUE),
          wentOutdoors = sum(break_condition_1, na.rm = TRUE),
          stretch = sum(break_condition_2, na.rm = TRUE),
          talked = sum(break_condition_4, na.rm = TRUE),
          other = sum(break_condition_8, na.rm = TRUE))

ba_2 = transpose(ba_1) %>%
  mutate(category = c( "junk", "Bathroom", "ST_School",
                        "ST_Socializing", "ST_Relax", "Went Outdoors",
                        "Moved/Stretched", "Talked/Socialized", "Other")) %>%
  select(category, V1, V2, V3) %>%
  filter(category != "junk") %>%
  mutate(usual = V1, outdoor = V2, movement = V3) %>%
  select(category, usual, outdoor, movement)

ba_3 = ba_2  %>% 
  pivot_longer(cols = c( "usual", "outdoor", "movement")) 

#write.csv(ba_3, "C:/Users/efrai/Documents/STAT_210_Spring_2025/efrain-ramirez-portfolio/portfolio-project/data/data_clean/break_activity_spent.csv")


ba_4 <- cbi24 %>%
  clean_names() %>%
  group_by(break_condition) %>%
  summarize(bathroom = sum(usual_break_3, na.rm = TRUE),
          screenSchool = sum(usual_break_5, na.rm = TRUE),
          screenSocial = sum(usual_break_6, na.rm = TRUE),
          screenRelax = sum(usual_break_7, na.rm = TRUE),
          wentOutdoors = sum(usual_break_1, na.rm = TRUE),
          stretch = sum(usual_break_2, na.rm = TRUE),
          talked = sum(usual_break_4, na.rm = TRUE),
          other = sum(usual_break_8, na.rm = TRUE))


ba_5 = transpose(ba_4) %>%
  mutate(category = c("junk", "Bathroom", "ST_School",
                        "ST_Socializing", "ST_Relax", "Went Outdoors",
                        "Moved/Stretched", "Talked/Socialized", "Other")) %>%
  select(category, V1, V2, V3 ) %>%
  filter(category != "junk" ) %>%
  mutate(usual = V1, outdoor = V2, movement = V3) %>%
  select(category, usual, outdoor, movement)

ba_6 = ba_5  %>% 
  pivot_longer(cols = c("usual", "outdoor", "movement")) 

#write.csv(ba_6, "C:/Users/efrai/Documents/STAT_210_Spring_2025/efrain-ramirez-portfolio/portfolio-project/data/data_clean/break_usual.csv")
```

```{r}
#| label: read-data
#read in your clean data here
break_activity_spent <- read_csv("data/data_clean/break_activity_spent.csv")
break_usual <- read_csv("data/data_clean/break_usual.csv")
```

## Visualzation 1

Here is a stacked bar plot with eight bars that each indicate the activity each participant did during their break separated by group break condition. 

```{r}
#| label: viz-1
#| fig-alt: "This is an alternative text for the first visualization"
#| fig-width: 6
#| fig-height: 4
#| fig-align: center

#create plot of iris dataset
break_activity_spent %>%
  ggplot( aes( x = category, y = value, fill = name  ) ) +
  geom_col( binwidth = 1, color = "grey", alpha = 0.5, position = "stack" ) + coord_flip() +
theme_minimal() +
  labs(x = "Break Activity", 
       y = "Number of Participants",
       title = "What did you do during break today(select all that apply)?",
       fill = "Break Condition") +
  theme(axis.title.y = element_text(angle = 0, vjust = 0.5)) +
  scale_fill_viridis_d()
```

## Visualization 2

Here is a stacked bar plot with eight bars that each indicate the activity each participant usually does during their break separated by group break condition. 

```{r}
#| label: viz-2
#| fig-alt: "This is an alternative text for the second visualization"
#| fig-width: 6
#| fig-height: 4
#| fig-align: center

#create plot of cars data set
break_usual %>%
  ggplot(aes( x = category, y = value, fill = name  ) ) +
  geom_col(binwidth = 1, color = "grey", alpha = 0.5, position = "stack") + coord_flip() +
theme_minimal() +
  labs(x = "Break Activity", 
       y = "Number of Participants",
       title = "How do you usually spend your break(select all that apply)?",
       fill = "Break Condition") +
  scale_fill_viridis_d()
```
